``Interestingness'' in Mathematical Theory Exploration

### 2019-11-29 PhDThesis/bucketing.tex
Include better bucketing graph

The pseudorandom and clustering implementations are more realistic, since they
do not use any information from the ground truth. These methods may also leave
some of their subsets empty (since, unlike the optimal and pessimal methods,
they are less prone to extremes).

@@

\begin{figure}
  % \input{images/bucketing}
  \includegraphics[width=\textwidth]{images/bucketing.png}
  \caption{Comparison of subset quality when selected by recurrent clustering
    (solid lines) versus pseudorandom selection (dashed lines). Darker lines
    show division into a greater number of subsets (between 1 and 20). The solid
    and dashed lines follow each other closely, indicating that clustering
    performs no better than random at signature selection.}
  \label{fig:bucketing}
\end{figure}

The signature selection problem is superficially similar to the
\emph{clustering} problem in machine learning: grouping data points based on
their similarity across multiple dimensions. To investigate this potential
relationship we created a signature selection method using the \emph{recurrent
  clustering} algorithm for Haskell code described in
\S~\ref{sec:recurrentclustering}: this flattens syntax trees into vectors,
substitutes fixed numbers for keywords and recurses to determine suitable
numbers for referenced expressions.

The results of this method are shown in Figure~\ref{fig:bucketing}, along with
the pseudorandom method as a control. The results of clustering turn out to be
no better than random; in fact the curves show surprisingly similar patterns.
\iffalse
TODO: Alison: Does this translate into a recommendation (e.g. not to
use clustering)? There should be some discussion of this - if not here then
point to another chapter where you discuss these results

>>>>

The pseudorandom and clustering implementations are more realistic, since they
do not use any information from the ground truth. We also permit these methods
to leave some of their subsets empty (since, unlike the optimal and pessimal
methods, they are less prone to extremes).

@@

\begin{figure}
  % \input{images/bucketed}
  \includegraphics[width=\textwidth]{images/bucketed.png}
  \caption{Quality evaluation of bucketing via recurrent clustering. Each plot
    uses a fixed number of buckets (5, 10, 15 or 20) and shows the number of
    extra theorems available when clusters are used as buckets, compared to a
    pseudorandom control. The difference is not statistically significant. There
    is a slight upward trend for larger sizes, but this is relatively small
    compared to the corresponding increase in total theorems available.}
  \label{fig:bucketing}
\end{figure}

The signature selection problem is superficially similar to the
\emph{clustering} problem in machine learning: grouping data points based on
their similarity across multiple dimensions. To investigate this potential
relationship we created a signature selection method using the \emph{recurrent
  clustering} algorithm for Haskell code described in
\S~\ref{sec:recurrentclustering}: this flattens syntax trees into vectors,
substitutes fixed numbers for keywords and recurses to determine suitable
numbers for referenced expressions. These vectors are clustered based on their
euclidean distance, using a \emph{k-means} algorithm, and these clusters are
chosen as our buckets.

The results of this method are shown in Figure~\ref{fig:bucketing}, which plots
the difference between bucketing based on this clustering approach, compared to
simply choosing buckets pseudorandomly (as a control). The results of clustering
are statistically indistinguishable from random.\footnote{Wilcoxon's signed-rank
  test finds no significant difference for most of the plotted points; whilst
  not a rigorous comparison (hence its ommission), this is nevertheless a clear
  indication of their similarity.}

\iffalse
TODO: Alison: Does this translate into a recommendation (e.g. not to
use clustering)? There should be some discussion of this - if not here then
point to another chapter where you discuss these results
